Undecidability and the structure of the Turing degrees

### 2018-08-15 paper.tex
Upper Semilattice result

\begin{lem}
  \label{lem:deg_count}
  Every Turing degree is of cardinality $\aleph_0$.
\end{lem}

@@

\begin{thm}
  $\mathcal{D}$ forms an upper semilattice.
  % Kleene-Post 54 Degrees upper semi-lattice, jump an operator on degrees
  \cite{kleene54_upper_semi_lattic_degrees_recur_unsol}
\end{thm}
\begin{lem}
 $ \mathcal{D}$ does not form a lattice.
  \cite{kleene54_upper_semi_lattic_degrees_recur_unsol}
\end{lem}

>>>>

\begin{lem}
  \label{lem:deg_count}
  Every Turing degree is of cardinality $\aleph_0$.
  \begin{proof}
    We have that since every Turing machine can be expressed as a finite string over
    $\{0,1\}$, it can be expressed as a finite subset of $\mathbb{N}$. So, it follows
    that for any given oracle, there are at most $\aleph_0$ oracle machines, so
    at most $\aleph_0$ problems are decidable for any Turing degree.
    \par
    Now that an upper bound has been established,  we show that the cardinality
    of each degree is exactly $\aleph_0$.\todo{}
  \end{proof}
\end{lem}

@@

\begin{thm}
  \label{thm:up_sem}
  $\mathcal{D}$ forms an upper semilattice.
  % Kleene-Post 54 Degrees upper semi-lattice, jump an operator on degrees
    \begin{proof}
    Observe that it is sufficient to prove that every two degrees have a
    supremum. Let $A,B\in \mathcal{D}$ and  $P_1\in A, P_2 \in B$. Consider the operation
    \[
      A\vee B = \deg\left( \{2n \mid n \in P_1\}\cup\{2n+1\cup n\in P_2\} \right),
    \]
    where $\deg(A)$ is the Turing degree of $A$. Now, we show that $A\vee B = \sup(A,B)$.\par First, we show that
    $A,B \leq_T A \vee B$. This is equivalent to showing
    that problems in $A$ and $B$ are decidable by oracle machines for $A\vee
    B$. But simple construction gives oracle machines that can decide $P_1,P_2$
    given an oracle for $A\vee B$: for any $n$, if $2n\in A\vee B$ then
    $n\in P_1$, and if $2n+1\in A\vee B$ then $n\in P_2$. This holds for any
    $P_1,P_2$, so it follows that $A,B \leq_T A \vee
    B$. \par Now, let $W\in \mathcal{D}$ such that $W \geq
    A,B$ and let $P\in A\vee B$. Then, we have that for every $n$, we can
    determine whether $n\in P$ by determining whether $n$ is even and
    $\frac{n}{2}\inP_1$ or odd and $\frac{n-1}{2}\in P_2$. Since both the parity
    of a number and $P_1$ and $P_2$ are decidable in $W$, it follows that
    $A\veeB$ is decidable in $W$, and thus $A\vee B \leq_T W$. So, it follows
    that $A\vee B = \sup(A,B)$. This holds for all $A,B\in\mathcal{D}$, so it
    follows that $\mathcal{D}$ forms an upper semilattice.  
  \end{proof}
\end{thm}

\begin{lem}
  \label{lem:not_lat}
  $ \mathcal{D}$ does not form a lattice.

\end{lem}
\begin{rem}
Theorem~\ref{thm:up_sem} and Lemma~\ref{lem:not_lat} were first shown in
Kleene's \emph{The Upper Semi-Lattice of Degrees of Recursive Unsolvability}~\cite{kleene54_upper_semi_lattic_degrees_recur_unsol}.
\end{rem}
